 
## Load and process the data
##### 1. Load the data 

```{r, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
library(data.table)
library(ggplot2)

fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl, destfile = paste0(getwd(), '/repdata%2Fdata%2Factivity.zip'), method = "curl")
unzip("repdata%2Fdata%2Factivity.zip",exdir = "data")
```


```{r, results='markup', warning=TRUE, message=TRUE}
activity_data <- data.table::fread(input = "data/activity.csv")
```

#What is mean total number of steps taken per day?
##1.Calculate the total number of steps taken per day

```{r}
Total_Steps <- activity_data[, c(lapply(.SD, sum, na.rm = FALSE)), .SDcols = c("steps"), by = .(date)] 

head(Total_Steps, 10)
```

## 1. Histogram of the total number of steps taken each day

```{r}
ggplot(Total_Steps, aes(x = steps)) +
    geom_histogram(fill = "blue", binwidth = 1000) +
    labs(title = "Daily Steps", x = "Steps", y = "Frequency")
```



## 2. Mean and median total number of steps taken per day

```{r}
Total_Steps[, .(Mean_Steps = mean(steps, na.rm = TRUE), Median_Steps = median(steps, na.rm = TRUE))]
```

-----

# Average daily activity pattern - steps pattern for each 5 minutes

```{r}
IntervalDT <- activity_data[, c(lapply(.SD, mean, na.rm = TRUE)), .SDcols = c("steps"), by = .(interval)] 

ggplot(IntervalDT, aes(x = interval , y = steps)) + geom_line(color="blue", size=1) + labs(title = "Avg. Daily Steps", x = "Interval", y = "Avg. Steps per day")
```

## 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps

```{r}
IntervalDT[steps == max(steps), .(max_interval = interval)]
```

----

# Missing values
## 1. The total number of missing values in the dataset 

```{r}
activity_data[is.na(steps), .N ]
```

### 2. Devise a strategy for filling in all of the missing values in the dataset.
### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
activity_data[is.na(steps), "steps"] <- activity_data[, c(lapply(.SD, median, na.rm = TRUE)), .SDcols = c("steps")]
```


##### 4. Histogram of the total number of steps taken each day 

```{r}
Total_Steps <- activity_data[, c(lapply(.SD, sum)), .SDcols = c("steps"), by = .(date)] 
ggplot(Total_Steps, aes(x = steps)) + geom_histogram(fill = "blue", binwidth = 1000) + labs(title = "Daily Steps", x = "Steps", y = "Frequency")
```


 

##### ... and Calculate and report the mean and median total number of steps taken per day. 

```{r}
Total_Steps[, .(Mean_Steps = mean(steps), Median_Steps = median(steps))]
```


----

# Are there differences in activity patterns between weekdays and weekends?
## 1. Create a new factor variable in the dataset with two levels for Weekday or Weekend

```{r}
activity_data$dateType <-  ifelse(as.POSIXlt(activity_data$date)$wday %in% c(0,6), 'weekend', 'weekday')
head(activity_data, 10)
```

## 2. Panel plot containing a time series plot

```{r}
activity_data[is.na(steps), "steps"] <- activity_data[, c(lapply(.SD, median, na.rm = TRUE)), .SDcols = c("steps")]
IntervalDT <- activity_data[, c(lapply(.SD, mean, na.rm = TRUE)), .SDcols = c("steps"), by = .(interval, `dateType`)] 

ggplot(IntervalDT , aes(x = interval , y = steps, color=`dateType`)) + geom_line() + labs(title = "Avg. Daily Steps by Weektype", x = "Interval", y = "No. of Steps") + facet_wrap(~`dateType` , ncol = 1, nrow=2)
```